---
title: "Object detection"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Multiple object classification 

The quiz assumes we've looked at the `classification_localization.Rmd` notebook and seen how to classify a single object in an image.


### Quiz 1

We've just seen how to classify a single object in an image. 
What if we wanted to classify multiple objects?

Choose all statements that are true.

```{r quiz1}
quiz(
  question("If we wanted to classify multiple objects in an image, ...",
    answer("... we would have to change the data generator to use multi-hot encoding", correct = TRUE),
    answer("we should use a sigmoid activation instead of softmax", correct = TRUE),
    answer("we should use mean squared error loss instead of categorical crossentropy"),
    answer("we should use binary crossentropy instead of categorical crossentropy", correct = TRUE)
  )
)
```


## Combining classification and regression

Think about constructing a model that does both classification and regression at the same time.

Check all that apply.

### Quiz 2


```{r quiz2}
quiz(
  question("To perform both classification and regression at the same time, we need to",
    answer("use the functional API with two inputs and two outputs"),
    answer("use the functional API with a single input and two outputs", correct = TRUE),
    answer("assign 4 units to the regression output and 1 unit to the class output"),
    answer("use the same loss function on both outputs"),
    answer("be prepared to experiment with scaling the loss functions, to give them equal / appropriate weights", correct = TRUE)
  )
)
```
