---
title: "Representation learning, interpretability, and uncertainty"
subtitle: "A pick of 'meta' topics in current DL"
author: "Sigrid Keydana"
date: "RStudio::conf 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
    css: xaringan-themer.css

---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_accent(
  base_color = "#006d6f"
)
```

```{r, load_refs, echo=FALSE, cache=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE, 
           bib.style = "authoryear", 
           cite.style = 'alphabetic', 
           style = "markdown",
           hyperlink = FALSE, 
           dashed = FALSE)
bib <- ReadBib("./bibliography.bib", check = FALSE)
```

# Representation learning, interpretability, and uncertainty

What do these topics have in common?

--

- Terms are not necessarily well-defined
- We'll keep out of terminology-related controversies, and
- Focus on the code

--

The idea being: Enable you to apply what's useful to you _in your area of work/interest_

---
class: inverse, middle, center

# Representation learning


---

# Why representation learning?

- Representation matters<sup>1</sup>
- Good representations make subsequent prediction task easier 
- Make use of abundant unlabeled data (unsupervised pretraining)
- Transfer learning / domain adaptation
- Density estimation
- Object generation


```{r, echo=FALSE, results=FALSE}
c1 <- Citet(bib, key = "Goodfellow-et-al-2016", .opts = list(cite.style = "authoryear"))
c2 <- Citet(bib, key = "5538", .opts = list(cite.style = "authoryear"))
```

.footnote[[1] cf. `r c1`, chap. 15, and `r c2`.]

---

## Variational autoencoders (VAEs)

- often used in object generation
- learn a _latent space_ that can be sampled from
- learning objective is a tradeoff between faithful input reconstruction and conforming to a (regularizing) prior
- Minimize: Evidence lower bound (ELBO)

$$ELBO\ = \ E[log\ p(x|z)]\ -\ KL(q(z)||p(z))$$

--

#### Problems with traditional VAEs

- may learn uninformative latent space
- tends to overfit the data

---

# InfoVAE and subtypes

- designed to learn more meaningful features
- e.g. by maximizing mutual information between latent space prior and posterior
- we'll see a subtype (MMD-VAE) soon

--

But before, we need some background on:

---
class: inverse, middle, center

# Eager execution


---

# Eager execution

- The non-graph way of doing TensorFlow/Keras
- More intuitive than static Keras/TF especially if not following the usual define-compile-fit cycle:
  - encoder-decoder models (as in GANs, VAEs ...)
  - neural style transfer
  - models with attention mechanisms
  
--

#### Detailed examples on the TensorFlow for R blog:

- [More flexible models with TensorFlow eager execution and Keras](https://blogs.rstudio.com/tensorflow/posts/2018-10-02-eager-wrapup/)
- [Image-to-image translation with pix2pix](https://blogs.rstudio.com/tensorflow/posts/2018-09-20-eager-pix2pix/)
- [Attention-based Image Captioning with Keras](https://blogs.rstudio.com/tensorflow/posts/2018-09-17-eager-captioning/)
- [Neural style transfer with eager execution and Keras](https://blogs.rstudio.com/tensorflow/posts/2018-09-10-eager-style-transfer/)
- [Generating images with Keras and TensorFlow eager execution](https://blogs.rstudio.com/tensorflow/posts/2018-08-26-eager-dcgan/)
- [Attention-based Neural Machine Translation with Keras](https://blogs.rstudio.com/tensorflow/posts/2018-07-30-attention-layer/)


---

# Eager execution closeup

- demo: [eager_intro.Rmd](../notebooks/eager_intro.Rmd)
- exercise: [4_meta_quizzes.Rmd](../exercises/4_meta_quizzes.Rmd)


---
class: inverse, middle, center

# Variational autoencoder (MMD-VAE)

---



---

# References

```{r, 'refs', results='asis', echo=FALSE}
PrintBibliography(bib)
```

